---
output: html_document
---
Analysis of Personal Movement Monitoring Data
========================
This is an analysis of personal movement monitoring data. 

**Loading the data**

```{r load}
repdata <-read.csv(unzip('C:/Documents and Settings/Jennifer/Desktop/Coursera/Repeatable Data/repdata-data-activity.zip'),header=TRUE)
summary(repdata)
```
**Mean Number of Total Steps**

First we total the steps per day and display the results in a histogram.  
```{r daytotals}
daytotals <- tapply(repdata$steps,INDEX=repdata$date,FUN=sum)
hist(daytotals,breaks=10)
```

We also calculate the mean and median of the number of steps taken per day.

```{r mean}
meanday <- mean(daytotals, na.rm=TRUE)
```
The mean is `r as.numeric(meanday)`.

```{r meadian}
medianday <- median(daytotals,na.rm=TRUE)
```
The median is `r medianday`.

**Average Daily Activity Pattern**

We average the number of steps taken for each 5-minute interval across all of the days and plot the result.

```{r intervalmeans}
intervalmeans <- aggregate(repdata$steps,by=list(repdata$interval),FUN=mean,na.rm=TRUE)

plot(intervalmeans,type="l",main="Average Daily Activity by Interval", ylab="Average Steps Taken", xlab="Interval Start Time")
```

We calculate which 5-minute interval contains the maximum number of steps.
```{r intervalmax}
max(intervalmeans$x)
intervalmax <- intervalmeans$Group.1[which.max(intervalmeans$x)]

```
The 5-minute interval containing the average maximum number of steps starts at `r intervalmax`.

**Inputing Missing Values**

First we calculate the total number of missing values (rows with NAs)
```{r missing}
library("plyr", lib.loc="C:/Program Files/R/R-3.1.2/library")
misses <- count(is.na(repdata$steps))[2,2]
```
The number of missing values (rows with NAs) is `r misses`.

Next we develop a strategy for replacing the missing values.  The strategy I chose was to replace the missing value with the mean for the 5-minute interval across all days. A new dataset is created using this strategy and the result is shown in a histogram.
```{r newdata}
newdata <- repdata
for (i in 1:17568)
     if (is.na(repdata$steps[i]))newdata$steps[i]=intervalmeans$x[which(intervalmeans$Group.1==repdata$interval[i])] else newdata$steps[i] = repdata$steps[i]
newtotals <- tapply(newdata$steps,INDEX=newdata$date,FUN=sum)
hist(newtotals,breaks=10)
newmean <- mean(newtotals)
print(newmean)
newmedian <- median(newtotals)
print(newmedian)
```

The effect of replacing the NAs with the means is to increase the means and to create 2 days which consist entirely of the mean values - as a result, the mean also becomes the median!
 


